Sex differences in upper limb musculoskeletal biomechanics during a lifting task

### 2020-03-25 paper/response/response.tex

% Preamble
\documentclass[preprint,review,12pt]{elsarticle}

% Package
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{lineno}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{mathpazo}
\usepackage{amsmath}
\usepackage{booktabs}

% Bibliography
\usepackage{natbib}
\bibliographystyle{elsarticle-num-names}

% Highlight
\usepackage{color,soul}
\DeclareRobustCommand{\hlcyan}[1]{{\sethlcolor{cyan}\hl{#1}}}

% Quote
\input{quote.tex}

% Document
\begin{document}

    \section*{Responses to Reviewers}
    \setcounter{section}{0}
    Dear Editor-in-Chief,

    Thank you for giving us the opportunity to submit a revised draft of our manuscript to Applied Ergonomics.
    We appreciate the time and effort that you, and the reviewers, have dedicated to providing your valuable feedback.
    We have incorporated changes to reflect the suggestions provided by the reviewers and have highlighted the changes within the manuscript.

    Here is a point-by-point response to the reviewers' comments and concerns.
    To simplify proofreading, we have quoted and underlined the text in yellow when an edit has been made to the original text.

    \section{Comments from Reviewer 1}\label{sec:reviewer-1}
    \begin{formal}
        The paper is very well structured;
        the figures are helpful to understand the results.
        The analysed parameters are interesting.
        The authors give supplementary information structured in Appendix.
    \end{formal}

    We would like to thank the reviewer for his/her time and insightful comments that will help us improve the article.

    \begin{formal}
        The parameter of the relative time spent beyond a shear-compression dislocation ratio by sex and mass (section 4.2. \textit{Sex and mass main effects}, figure 7) seems to be less representative for the conclusion.
    \end{formal}
    We agree that this parameter was less relevant to draw the conclusions of this study.
    We considered useful to include it because it links muscle activation, muscle forces and glenohumeral kinematics to the risk of upper limb injury, as stated in the \textit{Kinematic, electromyographic and musculoskeletal evidences of the sex-related differences} section:

    \begin{quote}
        "A greater muscle activation, muscle force and glenohumeral contribution during the handling task will indeed increase the solicitation of passive structures and alter the glenohumeral stability~\citep{Bergmann2007-zj}.
        This high ratio of shear and compression forces imposes high stresses on the stabilizing muscles and thus increases the risk of developing ULMD."
    \end{quote}

    \begin{formal}
        If we focus on the figure, approximatively 35\% (7 men subjects) are considered as outliers.
    \end{formal}

    Figure 7 is based on each of the 196 trials performed by the 20 women (98 trials) and 20 men (98 trials) --- and not on one aggregated data point per participant.
    The seven outliers that you describe are in fact seven trials which correspond to 7\% of the 98 trials performed by men:

    \begin{table}[H]
        \centering
        \caption{Outliers rate in Figure 7.}
        \begin{tabular}{lccc}
            \hline
            Sex & 6~kg & 12~kg & Total         \\ \hline
            Women & 1/98 (1.02\%) & 3/98 (3.06\%) & 4/196 (2.04\%) \\
            Men & 7/98 (7.14\%) & 7/98 (7.14\%) & 7/98 (7.14\%) \\ \hline
        \end{tabular}
    \end{table}

    We are sorry for this poor description.
    We added "for each trial" in Figure 7 caption to avoid confusion:

    \begin{quote}
        "Boxplot of the relative time spent beyond a shear-compression dislocation ratio \hl{for each trial} by sex (women in red and men in blue) and mass (6~kg on the left panel and 12~kg on the right panel) with median (horizontal lines), first-third interquartile range (bars), first-third interquartile range multiplied by 1.5 times the interquartile difference [$\textrm{Q1} - 1.5 \times \textrm{IQR}; \textrm{Q2} + 1.5 \times \textrm{IQR}$] (vertical lines).
        Data beyond the end of the whiskers are considered outliers (points)."
    \end{quote}

    \begin{formal}
        Despite this, the authors conclude "in addition, women spent more time beyond a shear-compression dislocation ratio." (section 5. \textit{Discussion}, page 15) and "we have also shown that women spend more time with a high risk of humerus dislocation." (section 5.2. \textit{Kinematic, electromyographic and musculoskeletal evidences of the sex-related differences}).
        More over, using a generic model, it is supposed that the "inter-individual variability is reduced" (section 5.3. \textit{Methodological considerations}).
        With these considerations, the authors are encouraged to be less affirmative on this point.
    \end{formal}

    While the outliers rate remains low, we fully agree and we have nuanced our interpretation of this parameter given the hypotheses it is based on.

    \textit{Abstract}:
    \begin{quote}
        "In addition, women \hl{might} spent more time beyond a shear-compression dislocation ratio."
    \end{quote}

    \textit{Highlights}, point 2:
    \begin{quote}
        "Women generated higher muscle forces and muscle activations when the box was above shoulder level and \hl{might} spent more time beyond a shear-compression dislocation ratio."
    \end{quote}

    \textit{Discussion}:
    \begin{quote}
        "In addition, women \hl{appear} to spend more time beyond a shear-compression dislocation ratio."
    \end{quote}

    \textit{Discussion, section 5.2. Kinematic, electromyographic and musculoskeletal evidences of the sex-related differences}:
    \begin{quote}
        "We have also shown that women \hl{might} spent more time with a high risk of humerus dislocation."
    \end{quote}

    \begin{formal}
        In terms of implications, the authors give different suggestions (section 5.4. \textit{Implications}) that are already known, for example "we recommend a technique that keeps the box closer from the trunk, reduces above shoulder work, makes greater use of the lower limbs and reduces the glenohumeral joint contribution".
        The authors are encouraged to be more ambitious and give some more insights on this point taking into account the results of the study.
    \end{formal}

    We agree that the recommendations could indeed rely more on the main finding of this study which is that women generated higher muscle forces and activations than men when the box was above shoulder level --- at either 6 or 12 kg.
    As such, we added the following sentences (\textit{Implications}, page 21):

    \begin{quote}
        \hl{"Taken together, our studies on kinematics~{\citep{Martinez2019-mm}}, \textsc{emg}~{\citep{Bouffard2019-fd}} and musculoskeletal biomechanics suggest that it is crucial for practitioners to give a special attention to women during above-shoulder work.
        Our indicators show that the biomechanical loads are higher in women compared to men while working above shoulder, which is critical for the development of shoulder injury~{\citep{Van_der_Molen2017-sb}}.
        It would therefore be necessary to adapt workload in women working in such position."}
    \end{quote}

    \section{Comment from Reviewer 2}\label{sec:reviewer-2}

    \begin{formal}
        Thank you for such a nice manuscript.
        You have made a great contribution to the investigation of sex differences in biomechanical loads at work.
    \end{formal}

    We thank the reviewer for his/her kind words and recommendation.

    \begin{formal}
        I only recommend you considering the s\textsc{emg} normalization also as a limitation, as we know that producing a real maximal contraction is something challenging.
        An underestimated MVC (particularly in women) can bias the results.
        If you have taken some special care on that issue I also recommend you reporting it in this study.
    \end{formal}

    While this study does not involve \textsc{emg} --- but rather muscle activations and forces computed by static optimization --- we agree that this is a limitation to be discussed for the results of our previous \textsc{emg} study~\citep{Bouffard2019-fd}, on which we rely to support and explain our results.
    Especially since it was not discussed in the original paper.
    \hlcyan{Since normalized \textsc{emg} is the product of the MVC and the measured \textsc{emg}, the MVC is indeed of key importance when comparing samples from different populations --- even more so in women, who may reach different levels of voluntary activation during MVC compared to men}~\citep{Ema2018-cd, Inglis2013-uq}.

    In \citet{Bouffard2019-fd}, the normalization procedure was as follows:
    \begin{quote}
        "A series of 12 submaximal voluntary contractions were performed to validate the electrode placement.
        Then, the same muscle contractions were performed with verbal encouragement twice at maximal voluntary intensity for normalisation purposes, in line with the recommendations of~\citet{Dal_Maso2016-wh}."
    \end{quote}

    While the "\textit{Methodological considerations}" section is limited to the specific limitations of this study, we added the following sentence in the previous section (\textit{Kinematic, electromyographic and musculoskeletal evidences of the sex-related differences}, page 17) when interpreting~\citet{Bouffard2019-fd} results:

    \begin{quote}
        \citet{Bouffard2019-fd} also reported higher muscle activation in women, particularly for prime movers.
        \hl{While both studies have specific limitations, they both point to similar conclusions.}
    \end{quote}

    The muscle activations reported in this study are also expressed in \%MVC\@.
    They actually depend on the muscle parameters implemented in the generic model we used.
    Although we tried to adjust the muscle parameters, this is an inherent limitation of using a generic model and one of the main limitations of our study, as stated in the "\textit{Methodological considerations}" section:

    \begin{quote}
        "In addition, the use of a generic model does not allow the musculoskeletal parameters to be personalized for each participant.
        As a result, inter-individual variability is reduced and the activation or strength of certain muscles may be overestimated."
    \end{quote}

    \bibliography{ref}

\end{document}

### 2020-03-25 paper/main.tex
Response to reviewer

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{lineno}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{natbib}
\usepackage{float}

@@

    % TODO
    %  - cover letter
    %  - fig

@@

        \begin{abstract}
            Women experience higher prevalence of work-related upper limb musculoskeletal disorders compared to men.
            Previous studies have investigated the biological, kinematic and electromyographic sex-related differences during a lifting task but the actual differences in musculoskeletal loads remain unknown.
            We investigated the sex differences in three musculoskeletal indicators: the sum of muscle activations, the sum of muscle forces and the relative time spent beyond a shear-compression dislocation ratio.
            A musculoskeletal model was scaled on 20 women and 20 men lifting a 6 or 12~kg box from hip to eye level.
            Women generated more muscle forces and activations than men, regardless of the lifted mass.
            Those differences occurred when the box was above shoulder level.
            In addition, women spent more time beyond a shear-compression dislocation ratio.
            Our work suggests higher musculoskeletal loads among women compared to men during a lifting task, which could be the result of poor technique and strength difference.
        \end{abstract}

@@

    \section{Highlights}\label{sec:highlights}
    \begin{itemize}
        \item We investigated sex differences in upper limb musculoskeletal biomechanics during a lifting task.
        \item Women generated higher muscle forces and muscle activations when the box was above shoulder level and spent more time beyond a shear-compression dislocation ratio.
        \item Those musculoskeletal indicators can be used to evaluate work technique and estimate the underlying musculoskeletal loads.
    \end{itemize}

>>>>

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{lineno}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{natbib}
\usepackage{float}
\usepackage{mathpazo}

@@

@@

        \begin{abstract}
            Women experience higher prevalence of work-related upper limb musculoskeletal disorders compared to men.
            Previous studies have investigated the biological, kinematic and electromyographic sex-related differences during a lifting task but the actual differences in musculoskeletal loads remain unknown.
            We investigated the sex differences in three musculoskeletal indicators: the sum of muscle activations, the sum of muscle forces and the relative time spent beyond a shear-compression dislocation ratio.
            A musculoskeletal model was scaled on 20 women and 20 men lifting a 6 or 12~kg box from hip to eye level.
            Women generated more muscle forces and activations than men, regardless of the lifted mass.
            Those differences occurred when the box was above shoulder level.
            In addition, women might spend more time beyond a shear-compression dislocation ratio.
            Our work suggests higher musculoskeletal loads among women compared to men during a lifting task, which could be the result of poor technique and strength difference.
        \end{abstract}

@@

    \section{Highlights}\label{sec:highlights}
    \begin{itemize}
        \item We investigated sex differences in upper limb musculoskeletal biomechanics during a lifting task.
        \item Women generated higher muscle forces and muscle activations when the box was above shoulder level and might spend more time beyond a shear-compression dislocation ratio.
        \item Those musculoskeletal indicators can be used to evaluate work technique and estimate the underlying musculoskeletal loads.
    \end{itemize}

### 2020-03-25 paper/parts/3-results.tex
Response to reviewer

The distribution of muscle activations (Figure~\ref{fig:dist}, left panel) reveals that several muscles are weakly activated (100\% of the time $<$20\%~MVC), particularly the \textsc{tmaj}, \textsc{tric}, \textsc{pecm}3, \textsc{delt}3, \textsc{pecm}2, \textsc{sbcl}, \textsc{corb}, \textsc{trp}4, \textsc{pmn}.
The five most activated muscles are \textsc{trp}1, \textsc{infsp}, \textsc{delt}1, \textsc{delt}2 and \textsc{trp}2--each with a large activation range (10-80\% MVC).
Similarly, the distribution of muscles forces (Figure~\ref{fig:dist}, right panel) reveals that many muscles are lightly involved (100\% of the time $<$100~N) (\textsc{tmaj}, \textsc{rmj}2, \textsc{sbcl}, \textsc{pecm}3, \textsc{rmj}1, \textsc{pmn}, \textsc{corb}, \textsc{pecm}2, \textsc{delt}3 and \textsc{trp}3) and a similar muscles group with high forces (\textsc{infsp}, \textsc{delt}2, \textsc{delt}1, \textsc{lat}, \textsc{trp}1).

@@

\begin{figure}[H]
    \centering\includegraphics[width=0.3\linewidth]{fig/dislocation.pdf}
    \caption{Boxplot of the relative time spent beyond a shear-compression dislocation ratio by sex (women in red and men in blue) and mass (6~kg on the left panel and 12~kg on the right panel) with median (horizontal lines), first-third interquartile range (bars), first-third interquartile range multiplied by 1.5 times the interquartile difference [$\textrm{Q1} - 1.5 \times \textrm{IQR}; \textrm{Q2} + 1.5 \times \textrm{IQR}$] (vertical lines).
    Data beyond the end of the whiskers are considered outliers (points).}
    \label{fig:dislocation}
\end{figure}

>>>>

The distribution of muscle activations (Figure~\ref{fig:dist}, left panel) reveals that several muscles are weakly activated (100\% of the time $<$20\%~MVC), particularly the \textsc{tmaj}, \textsc{tric}, \textsc{pecm}3, \textsc{delt}3, \textsc{pecm}2, \textsc{sbcl}, \textsc{corb}, \textsc{trp}4, \textsc{pmn}.
The five most activated muscles are \textsc{trp}1, \textsc{infsp}, \textsc{delt}1, \textsc{delt}2 and \textsc{trp}2--each with a large activation range (10-80\% MVC).
Similarly, the distribution of muscle forces (Figure~\ref{fig:dist}, right panel) reveals that many muscles are lightly involved (100\% of the time $<$100~N) (\textsc{tmaj}, \textsc{rmj}2, \textsc{sbcl}, \textsc{pecm}3, \textsc{rmj}1, \textsc{pmn}, \textsc{corb}, \textsc{pecm}2, \textsc{delt}3 and \textsc{trp}3) and a similar muscles group with high forces (\textsc{infsp}, \textsc{delt}2, \textsc{delt}1, \textsc{lat}, \textsc{trp}1).

@@

\begin{figure}[H]
    \centering\includegraphics[width=0.3\linewidth]{fig/dislocation.pdf}
    \caption{Boxplot of the relative time spent beyond a shear-compression dislocation ratio for each trial by sex (women in red and men in blue) and mass (6~kg on the left panel and 12~kg on the right panel) with median (horizontal lines), first-third interquartile range (bars), first-third interquartile range multiplied by 1.5 times the interquartile difference [$\textrm{Q1} - 1.5 \times \textrm{IQR}; \textrm{Q2} + 1.5 \times \textrm{IQR}$] (vertical lines).
    Data beyond the end of the whiskers are considered outliers (points).}
    \label{fig:dislocation}
\end{figure}

### 2020-03-25 paper/parts/4-discussion.tex
Response to reviewer

In this study, we investigated the sex and mass effects on three musculoskeletal indicators when women and men lifted 6~kg and 12~kg boxes from hip to eye level.
As hypothesized, women generated more muscle force and activation than men, regardless of the mass lifted.
Those differences occurred during the dropping phase, when the box was above shoulder level.
In addition, women spent more time beyond a shear-compression dislocation ratio.

@@

It seems unlikely that the kinematic alterations just described do not lead to a different muscle coordination scheme.
During the experimental task, women indeed generated a higher sum of muscle activations and forces than men during the dropping phase.
\citet{Bouffard2019-fd} also reported higher muscle activation in women, particularly for prime movers.
In both studies, the muscle activation in the 90\textsuperscript{th} percentile reached by women seems to be a major muscular effort to be maintained in professional settings (around 65\% MVC in women against 50\% MVC in men).

We have also shown that women spend more time with a high risk of humerus dislocation.
A greater muscle activation, muscle force and glenohumeral contribution during the handling task will indeed increase the sollication of passive structures and alter the glenohumeral stability~\cite{Bergmann2007-zj}.
This high ratio of shear and compression forces imposes high stresses on the stabilizing muscles and thus increases the risk of developing \textsc{ulmd}.
The infraspinatus, anterior deltoid and lateral deltoid muscles are all three stabilizers of the humerus~\cite{Blache2017-pv, Yanagawa2008-es} and are also the muscles with the highest sum of muscle forces in our dataset.
This underlines the importance of strengthening those muscles for manual handling workers to maintain shoulder stability and avoid injuries~\cite{Sharkey1995-gl}.

@@

Work-related musculoskeletal injury is a complex multi-causal phenomenon.
Our results suggest a careful consideration of sex during ergonomic interventions on overhead tasks.
With our work on kinematics, electromyography and musculoskeletal modelling, we developed new biomechanical \textsc{ulmd} risk indicators.
The development of synthetic indicators is a step towards objective quantification of exposure to physical risk factors.
They can be used to evaluate a particular work task or technique and estimate the underlying musculoskeletal loads.
They are, for example, already applied in ergonomics investigation to report the expertise-related difference during a lifting task \cite{goubault-evsn}.
In our case, these synthetic indicators emphasized the importance of a proper lifting technique on musculoskeletal loads and by extension \textsc{ulmd} prevalence.
In order to mitigate the musculoskeletal loads on the upper limb during a lifting task, we recommend a technique that keeps the box closer from the trunk, reduces above shoulder work, makes greater use of the lower limbs and reduces the glenohumeral joint contribution.
Practitioners may also consider adapting the work environment to include lighter boxes, stabilizing muscle strengthening sessions and more generally physical activity to improve muscle strength and endurance.
These incentives are particularly recommended for female workers to compensate for the strength differential compared to men.
Beyond sex-related differences, we defend an ergonomic practice based on a process tailored to each worker and based on multidisciplinary scientific works.

>>>>

In this study, we investigated the sex and mass effects on three musculoskeletal indicators when women and men lifted 6~kg and 12~kg boxes from hip to eye level.
As hypothesized, women generated more muscle force and activation than men, regardless of the mass lifted.
Those differences occurred during the dropping phase, when the box was above shoulder level.
In addition, women appear to spent more time beyond a shear-compression dislocation ratio.

@@

It seems unlikely that the kinematic alterations just described do not lead to a different muscle coordination scheme.
During the experimental task, women indeed generated a higher sum of muscle activations and forces than men during the dropping phase.
\citet{Bouffard2019-fd} also reported higher muscle activation in women, particularly for prime movers.
While both studies have specific limitations, they both point to similar conclusions.
In both studies, the muscle activation in the 90\textsuperscript{th} percentile reached by women seems to be a major muscular effort to be maintained in professional settings (around 65\% MVC in women against 50\% MVC in men).

We have also shown that women might spend more time with a high risk of humerus dislocation.
A greater muscle activation, muscle force and glenohumeral contribution during the handling task will indeed increase the solicitation of passive structures and alter the glenohumeral stability~\cite{Bergmann2007-zj}.
This high ratio of shear and compression forces imposes high stresses on the stabilizing muscles and thus increases the risk of developing \textsc{ulmd}.
The infraspinatus, anterior deltoid and lateral deltoid muscles are all three stabilizers of the humerus~\cite{Blache2017-pv, Yanagawa2008-es} and are also the muscles with the highest sum of muscle forces in our dataset.
This underlines the importance of strengthening those muscles for manual handling workers to maintain shoulder stability and avoid injuries~\cite{Sharkey1995-gl}.

@@

Work-related musculoskeletal injury is a complex multi-causal phenomenon.
Our results suggest a careful consideration of sex during ergonomic interventions on overhead tasks.
With our work on kinematics, electromyography and musculoskeletal modelling, we developed new biomechanical \textsc{ulmd} risk indicators.
The development of synthetic indicators is a step towards objective quantification of exposure to physical risk factors.
They can be used to evaluate a particular work task or technique and estimate the underlying musculoskeletal loads.
They are, for example, already applied in ergonomics investigation to report the expertise-related difference during a lifting task \cite{goubault-evsn}.
In our case, these synthetic indicators emphasized the importance of a proper lifting technique on musculoskeletal loads and by extension \textsc{ulmd} prevalence.
In order to mitigate the musculoskeletal loads on the upper limb during a lifting task, we recommend a technique that keeps the box closer from the trunk, reduces above shoulder work, makes greater use of the lower limbs and reduces the glenohumeral joint contribution.
Practitioners may also consider adapting the work environment to include lighter boxes, stabilizing muscle strengthening sessions and more generally physical activity to improve muscle strength and endurance.
These incentives are particularly recommended for female workers to compensate for the strength differential compared to men.
Taken together, our studies on kinematics~\citep{Martinez2019-mm}, \textsc{emg}~\citep{Bouffard2019-fd} and musculoskeletal biomechanics suggest that it is crucial for practitioners to give a special attention to women during above-shoulder work.
Our indicators show that the biomechanical loads are higher in women compared to men while working above shoulder, which is critical for the development of shoulder injury~\citep{Van_der_Molen2017-sb}.
It would therefore be necessary to adapt workload in women working in such position.
Beyond sex-related differences, we defend an ergonomic practice based on a process tailored to each worker and based on multidisciplinary scientific works.
